---
title: "ABC Phenomics Supplemental Data Analysis"
author: "Sophie Watts"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    code_folding: hide
    theme: cerulean
    toc: yes
    toc_float: yes
---
```{r setup, include=FALSE}
#load packages
library("readr")
library("dplyr")
library("RColorBrewer")
library("gplots")
library("viridis")
library("scales")
library("tidyverse")
library("cowplot")
library("gridExtra")
library("grid")
library("plotrix")

theme_set(theme_classic())

```

#Introduction
This script contains the code to create the supplemental tables and figures for the the ABC Phenomics Manuscript.

#Supp Figure 1: Orchard map.

#Supp Figure 2: Correlation plot of TPC vs. FRAP

```{r}
#load phenotype data.
abc_phenos <- read_csv("../outputs/pheno_meta_data.csv")

#select phenolics data.
abc_phenos <- select(abc_phenos, tpc, frap)

supp_figure_2 <- ggplot(abc_phenos, aes(tpc, frap))+
  geom_point(size=1, alpha=0.7, colour = "darkgrey")+
  theme_bw()+
  labs(x="Total phenolic content", y="Ferric reducing antioxidant powers")+
  theme(panel.border = element_blank(), axis.text=element_text(colour = "black", size=10), axis.title=element_text(size = 10, face = "bold"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black", size = 0.2), plot.title = element_text(size = 9))+
  coord_fixed(xlim = c(0,40), ylim= c(0,40))+
  stat_smooth(method="lm", col="black", se=FALSE)+
  ggtitle(paste("r2 = ", signif(cor.test(abc_phenos$tpc, abc_phenos$frap, method = "pearson")$estimate^2, digits = 2), " p-value =", signif(cor.test(abc_phenos$tpc, abc_phenos$frap, method = "pearson")$p.value, digits = 2)))
ggsave("supp_figure_2", path = "../figures_tables/supp_figures", device = "pdf")

```
#Supp. Figure 3: Correlation fruit quality traits between 2016 and 2017.
```{r}

abc_phenos <- read_csv("../outputs/pheno_meta_data.csv")

#load in test table with new names of the phenos that are shared between the two years and this table has been made with new pheno names so that we can more easily run a loop.
name_table <- read_delim("../data/2016_2017_pheno_table.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

abc_matrix = as.matrix(abc_phenos)

pdf("supp_figure_3.pdf", width = 10, height = 12)
par(mfrow = c(5, 4))
for (i in 1:nrow(name_table)) {
  names_vec = as.character(name_table[i,])
  pheno1 = as.numeric(abc_matrix[,names_vec[1]])
  pheno2 = as.numeric(abc_matrix[,names_vec[2]])
  index_both_phenos = which(!is.na(pheno1) & !is.na(pheno2))
  maxlim = max(c(pheno1[index_both_phenos],pheno2[index_both_phenos]),na.rm = T)
  maxlim = maxlim * 1.1
  minlim = min(c(pheno1[index_both_phenos],pheno2[index_both_phenos]),na.rm = T)
  minlim = minlim * 0.9
  cortest = cor.test(pheno1, pheno2, method = "pearson")
  r2 = as.numeric(round(cortest$estimate^2, 3))
  pval = as.numeric(signif(cortest$p.val, 3))*17
  plot(pheno1, pheno2, pch=16, col="darkgrey", xlab = "2016", ylab = "2017", sub = paste("r2 = ", r2, " p =", pval, " N =", length(index_both_phenos)), main = paste(names_vec[3], names_vec[4]), xlim = c(minlim, maxlim), ylim = c(minlim, maxlim), abline(lm(pheno2 ~ pheno1)))
}
dev.off()



```
#Supp. Figure 4: Histograms of phenotypes from 2016.
```{r}
abc_phenos <- read_csv("../outputs/pheno_meta_data.csv")

abc_phenos <- abc_phenos %>% select(apple_id, date_jul_16_harv, time_ripen_16_harv, weight_avg_16_harv, firmness_avg_16_harv, brix_avg_16_harv, acidity_16_harv)  %>% gather(key="pheno", value="value", -apple_id) 

#change the phenotype to a factor and reorder the factor
class(abc_phenos$pheno)
abc_phenos$pheno <- as.factor(abc_phenos$pheno)
levels(abc_phenos$pheno)

#set the levels for the factors and reoder the levels/rename
abc_phenos$pheno <- factor(abc_phenos$pheno, levels=c("date_jul_16_harv", "time_ripen_16_harv", "weight_avg_16_harv", "firmness_avg_16_harv", "brix_avg_16_harv", "acidity_16_harv"))
levels(abc_phenos$pheno)


#filter the top ten apples then calculate the max and the median, then summarize to get min and max
top_ten <-  abc_phenos %>% filter(apple_id %in% c(48, 205, 160, 167, 1223, 222, 57, 235, 171))

#caluclate min of each trait
min_top_ten <- top_ten %>% group_by(pheno) %>% slice(which.min(value))

#caluclate max of each trait
max_top_ten <- top_ten %>% group_by(pheno) %>% slice(which.max(value))

#calculate the median of the top ten
median_top_ten <- top_ten %>% group_by(pheno) %>% summarise(median = median(value, na.rm=T))

#calculate IQR
iqr_top_ten <- top_ten %>% group_by(pheno) %>% summarise(iqr_top_ten= IQR(value, na.rm=T))

#make a table with the max and min values together
min_max_top_ten <- left_join(min_top_ten, max_top_ten, by = "pheno")
min_max_top_ten <- left_join(min_max_top_ten, median_top_ten, by = "pheno")
min_max_top_ten <- left_join(min_max_top_ten, iqr_top_ten, by = "pheno")

min_max_top_ten <- min_max_top_ten %>% select(-c(apple_id.x, apple_id.y))

#rename columns
colnames(min_max_top_ten)[colnames(min_max_top_ten)=="value.x"] <- "min_top"
colnames(min_max_top_ten)[colnames(min_max_top_ten)=="value.y"] <- "max_top"
colnames(min_max_top_ten)[colnames(min_max_top_ten)=="median"] <- "median_top"


min_max_top_ten <- min_max_top_ten %>% group_by(pheno)

#make labels for each panel in the facet grid
labels <- c(date_jul_16_harv = "Harvest date (Julian days) N=608", time_ripen_16_harv ="Time to ripen (julian days) N=607", weight_avg_16_harv = "Weight (g) N=609", firmness_avg_16_harv = "Firmness (kg/cm) N=610", brix_avg_16_harv = "Soluble solids (Brix) N=608", acidity_16_harv = "Acidity (g/L) N=571")

#make histograms for supp fig 4
pdf("supp_figure_4.pdf", width = 9, height = 6)
abc_phenos %>% ggplot(aes(x=value))+
  geom_rect(data = min_max_top_ten, aes(ymin = 0, ymax = Inf, xmin = min_top, xmax = max_top), fill = "lightgrey", inherit.aes = FALSE)+
  geom_histogram(bins=30, size = 0.8)+
  ylab("Count")+
  scale_x_continuous(expand = c(0,0))+
  scale_y_continuous(expand = c(0,0))+
  facet_wrap(~pheno, scale='free', nrow=2, strip.position = "bottom", labeller=labeller(pheno = labels))+
  theme(strip.placement = "outside", strip.background = element_blank(), panel.spacing = unit(2, "lines"), axis.title.x = element_blank(), strip.text.x = element_text(size = 8, face = "bold"), axis.text.x = element_text(size = 7), axis.text.y = element_text(size = 7))
dev.off()




```
#Supp. Figure 5: 2016 storage phenotypes
```{r}
abc_phenos <- read_csv("../outputs/pheno_meta_data.csv")


acid_percent <- abc_phenos %>% select(apple_id, percent_acidity_16) %>% gather(key="pheno", value="value", -apple_id) 
top_ten_acid_percent <-  acid_percent %>% filter(apple_id %in% c(48, 205, 160, 167, 1223, 222, 57, 235, 171))


brix_percent <- abc_phenos %>% select(apple_id, percent_brix_16) %>% gather(key="pheno", value="value", -apple_id) 
top_ten_brix_percent <-  brix_percent %>% filter(apple_id %in% c(48, 205, 160, 167, 1223, 222, 57, 235, 171))


firmness_percent <- abc_phenos %>% select(apple_id, percent_firmness_avg_16) %>% gather(key="pheno", value="value", -apple_id) 
top_ten_firmness_percent <-  firmness_percent %>% filter(apple_id %in% c(48, 205, 160, 167, 1223, 222, 57, 235, 171))


weight_percent <- abc_phenos %>% select(apple_id, percent_weight_16) %>% gather(key="pheno", value="value", -apple_id)
top_ten_weight_percent <-  weight_percent %>% filter(apple_id %in% c(48, 205, 160, 167, 1223, 222, 57, 235, 171))


pdf("supp_figure_5.pdf", width = 8, height = 10)

a1 <- ggplot(abc_phenos, aes(acidity_16_harv, acidity_16_stor))+
  geom_point(size=1, alpha=0.7, colour = "darkgrey")+
  theme_bw()+
  labs(x="Acidity at harvest (g/L)", y="Acidity after storage (g/L)")+
  theme(panel.border = element_blank(), axis.text=element_text(colour = "black", size=10), axis.title=element_text(size = 10, face = "bold"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black", size = 0.2), plot.title = element_text(size = 9))+
  coord_fixed(xlim = c(0,18), ylim= c(0,18))+
  stat_smooth(method="lm", col="black", se=FALSE)+
  ggtitle(paste("r2 = ", signif(cor.test(abc_phenos$acidity_16_harv, abc_phenos$acidity_16_stor, method = "pearson")$estimate^2, digits = 2), " p-value =", signif(cor.test(abc_phenos$acidity_16_harv, abc_phenos$acidity_16_stor, method = "pearson")$p.value, digits = 2)))


a2 = ggplot(data=abc_phenos, aes(abc_phenos$percent_acidity_16))+
  geom_rect(data = top_ten_acid_percent, aes(ymin = 0, ymax = Inf, xmin = min(value, na.rm = T), xmax = max(value, na.rm = T)), fill = "lightgrey", inherit.aes = FALSE)+
  geom_histogram(bins = 50)+
  scale_x_continuous(name = "Change in Acidity (%) N=333")+
  scale_y_continuous(name = "Count", expand = c(0,0))+
  theme(axis.text=element_text(colour = "black", size=9), axis.title=element_text(size = 10, face = "bold"))


##################################

b1 <- ggplot(abc_phenos, aes(brix_avg_16_harv, brix_16_stor))+
  geom_point(size=1, alpha=0.7, colour = "darkgrey")+
  theme_bw()+
  labs(x="SSC at harvest (Brix)", y="SSC after storage (Brix)")+
  theme(panel.border = element_blank(), axis.text=element_text(colour = "black", size=10), axis.title=element_text(size = 10, face = "bold"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black", size = 0.2), plot.title = element_text(size = 9))+
  coord_fixed(xlim = c(7,21), ylim= c(7,21))+
  stat_smooth(method="lm", col="black", se = FALSE)+
  ggtitle(paste("r2 = ", signif(cor.test(abc_phenos$brix_avg_16_harv, abc_phenos$brix_16_stor, method = "pearson")$estimate^2, digits = 2), " p-value =", signif(cor.test(abc_phenos$brix_avg_16_harv, abc_phenos$brix_16_stor, method = "pearson")$p.value, digits = 2)))
  
  

b2 = ggplot(data=abc_phenos, aes(abc_phenos$percent_brix_16))+
  geom_rect(data = top_ten_brix_percent, aes(ymin = 0, ymax = Inf, xmin = min(value, na.rm = T), xmax = max(value, na.rm = T)), fill = "lightgrey", inherit.aes = FALSE)+
  geom_histogram(bins = 50)+
  scale_x_continuous(name = "Change in SSC (%) N=343")+
  scale_y_continuous(name = "Count", expand = c(0,0))+
  theme(axis.text=element_text(colour = "black", size=9), axis.title=element_text(size = 10, face = "bold"))

#############################################

f1 <- ggplot(abc_phenos, aes(firmness_avg_16_harv, firmness_avg_16_stor))+
  geom_point(size=1, alpha=0.7, colour = "darkgrey")+
  theme_bw()+
  labs(x="Firmness at harvest (kg/cm2)", y="Firmness after storage (kg/cm2)")+
  theme(panel.border = element_blank(), axis.text=element_text(colour = "black", size=10), axis.title=element_text(size = 10, face = "bold"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black", size = 0.2), plot.title = element_text(size = 9))+
  coord_fixed(xlim = c(2,14), ylim= c(2,14))+
  stat_smooth(method="lm", col="black", se = FALSE)+
  ggtitle(paste("r2 = ", signif(cor.test(abc_phenos$firmness_avg_16_harv, abc_phenos$firmness_avg_16_stor, method = "pearson")$estimate^2, digits = 3), " p-value = ", signif(cor.test(abc_phenos$firmness_avg_16_harv, abc_phenos$firmness_avg_16_stor, method = "pearson")$p.value, digits = 2)))


f2 = ggplot(data=abc_phenos, aes(abc_phenos$percent_firmness_avg_16))+
  geom_rect(data = top_ten_firmness_percent, aes(ymin = 0, ymax = Inf, xmin = min(value, na.rm = T), xmax = max(value, na.rm = T)), fill = "lightgrey", inherit.aes = FALSE)+
  geom_histogram(bins = 50)+
  scale_x_continuous(name = "Change in Firmness (%) N=352")+
  scale_y_continuous(name = "Count", expand = c(0,0))+
  theme(axis.text=element_text(colour = "black", size=9), axis.title=element_text(size = 10, face = "bold"))


#############################################


w1 <- ggplot(abc_phenos, aes(weight_avg_16_harv, weight_avg_16_stor))+
  geom_point(size=1, alpha=0.7, colour = "darkgrey")+
  theme_bw()+
  labs(x="Weight at harvest (g)", y="Weight after storage (g)")+
  theme(panel.border = element_blank(), axis.text=element_text(colour = "black", size=10), axis.title=element_text(size = 10, face = "bold"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black", size = 0.2), plot.title = element_text(size = 9))+
  coord_fixed(xlim = c(42,350), ylim= c(42,350))+
  stat_smooth(method="lm", col="black", se = FALSE)+
  ggtitle(paste("r2 = ", signif(cor.test(abc_phenos$weight_avg_16_harv, abc_phenos$weight_avg_16_stor, method = "pearson")$estimate^2, digits = 3), " p-value = ", signif(cor.test(abc_phenos$weight_avg_16_harv, abc_phenos$weight_avg_16_stor, method = "pearson")$p.value, digits = 2)))  



w2 = ggplot(data=abc_phenos, aes(abc_phenos$percent_weight_16))+
  geom_rect(data = top_ten_weight_percent, aes(ymin = 0, ymax = Inf, xmin = min(value, na.rm = T), xmax = max(value, na.rm = T)), fill = "lightgrey", inherit.aes = FALSE)+
  geom_histogram(bins = 50)+
  scale_x_continuous(name = "Change in Weight (%) N=355")+
  scale_y_continuous(name = "Count", expand = c(0,0))+
  theme(axis.text=element_text(colour = "black", size=9), axis.title=element_text(size = 10, face = "bold"))


list_of_figures <- list(a1, a2, b1, b2, f1, f2, w1, w2)

cowplot::plot_grid(
  plotlist = list_of_figures,
  nrow = 4, 
  align = 'hv',
  labels = "auto",
  hjust = -1
)
dev.off()

```
#Supp. figure 6: 2016 heatmap.
```{r}

#make supplemental figure 5 heatmap of correlations between phenotypes for 2016 data.
#Load phenotype table
abc_phenos <- read_csv("../outputs/pheno_meta_data.csv")

abc_phenos <- abc_phenos %>% select(precocity_16_harv, flowering_jul_16_harv, time_ripen_16_harv, date_jul_16_harv, juiciness_16_harv, weight_avg_16_harv, firmness_avg_16_harv, brix_avg_16_harv, acidity_16_harv, brix_acid_16_harv, percent_weight_16, percent_firmness_avg_16, percent_brix_16, percent_acidity_16)

abc_phenos[1:14] <- lapply(abc_phenos[1:14], as.character)
abc_phenos[1:14] <- lapply(abc_phenos[1:14], as.numeric)

#change from tbl to base dataframe.
abc_phenos <- as.data.frame(abc_phenos)

#create an empty matrix for correlations, make matrix with no data and just col and rows from final pheno table.
pairwise_pheno_correlations=matrix(,ncol(abc_phenos), ncol(abc_phenos))
rownames(pairwise_pheno_correlations)=colnames(abc_phenos)
colnames(pairwise_pheno_correlations)=colnames(abc_phenos)
#marix for pvalues.
pairwise_pheno_correlations_pval=matrix(,ncol(abc_phenos), ncol(abc_phenos))
rownames(pairwise_pheno_correlations_pval)=colnames(abc_phenos)
colnames(pairwise_pheno_correlations_pval)=colnames(abc_phenos)

#Treat the data as all quantitative data and run pearson's correlation.
for (i in 1:ncol(abc_phenos)) {
  phenoname_x = colnames(abc_phenos)[i]
  for (j in 1:ncol(abc_phenos)) {
    phenoname_y = colnames(abc_phenos)[j]
    pairwise_pheno_correlations[j,i]=cor.test(abc_phenos[,i], abc_phenos[,j], method = "pearson")$estimate^2
    pairwise_pheno_correlations_pval[j,i]= cor.test(abc_phenos[,i], abc_phenos[,j], method = "pearson")$p.value
  }
}



#bonneferonni correct
pairwise_pheno_correlations_pval[upper.tri(pairwise_pheno_correlations_pval)] = p.adjust(pairwise_pheno_correlations_pval[upper.tri(pairwise_pheno_correlations_pval)], method = "bonferroni")
pairwise_pheno_correlations_pval[lower.tri(pairwise_pheno_correlations_pval)] = p.adjust(pairwise_pheno_correlations_pval[lower.tri(pairwise_pheno_correlations_pval)], method = "bonferroni")



#make the heat map/
#Set colour.
hmcol = colorRampPalette(brewer.pal(9, "BrBG"))

#Get one half of correlation square (aka one triangle). Set lower to NA.
pairwise_pheno_correlations[lower.tri(pairwise_pheno_correlations)] <- NA

#make heat map for correlations without key.
pdf("supp_figure_6_heatmap.pdf", width = 8, height = 8)
mar.default <- c(0,0,0,0)
par(mar=mar.default, cex = 0.5)
heatmap.2(pairwise_pheno_correlations,
          col = hmcol(1000),
          key=T,
          keysize = 2.5,
          symm = T,
          Colv = F,
          Rowv = F,
          trace="none",
          dendrogram = "none",
          density.info = "none",
          symbreaks = T,
          colsep = c(0:14),
          rowsep = c(0:14),
          sepcolor = "white",
          sepwidth = c(0.05, 0.05),
          symkey = T,
          margin = c(12,12)
          )
dev.off()

#heat map for pvals of correlations.

#Get one half of correltaion square (aka one triangle). Set lower to NA.
pairwise_pheno_correlations_pval[upper.tri(pairwise_pheno_correlations_pval)] <- NA

#Pick colours.
hmcol_p = rev(brewer.pal(3, "YlGn"))
hmcol_p[3] = "#F7F7F7"

pdf("supp_figure_6_heatmap_pval.pdf", width = 8, height = 8)
mar.default <- c(0,0,0,0)
par(mar=mar.default, cex = 0.5)
heatmap.2(pairwise_pheno_correlations_pval,
          col = hmcol_p,
          key=T,
          keysize = 0.5,
          symm = T,
          Colv = F,
          Rowv = F,
          trace="none",
          dendrogram = "none",
          density.info = "none",
          breaks = c(0, 0.01, 0.05, 1),
          revC = F,
          symbreaks = T,
          colsep = c(0:28),
          rowsep = c(0:28),
          sepcolor = "white",
          sepwidth = c(0.05, 0.05),
          symkey = T,
          margin = c(12, 12)
          )
dev.off()

#save p-values and rhos for supplemental table.
write.table(pairwise_pheno_correlations, file = "../figures_tables/supp_tables/supp_table_5_r2.csv", quote = F, sep = ",", row.names = F)

write.table(pairwise_pheno_correlations_pval, file = "../figures_tables/supp_tables/supp_table_5_pvals.csv", quote = F, sep = ",", row.names = F)

```
#Supp. Table 1: This table is pheno_meta_data.csv

#Supp. Table 2: This is a data dictionary that explains the column names of Supp Table 1 and how they were measured.

#Supp. Table 3: Table showing the maximum, minimum and median values of fruit quality traits from the Top 9 and the entire ABC, as well as the wilcoxin p-values to detemrine whether the two groups differed significantly.
```{r}
#make supplemental table 3 of min, max and median.

#Load phenotype table
abc_phenos <- read_csv("../outputs/pheno_meta_data.csv")

abc_pheno_field <- abc_phenos %>% select(apple_id, precocity_16_harv, flowering_jul_16_harv, time_ripen_2017, date_jul_17_harv, juiciness_16_harv, weight_avg_17_harv, firmness_avg_17_harv, brix_17_harv, acidity_17_harv, tpc, percent_weight_avg_17, percent_firmness_avg_17, percent_brix_17, percent_acidity_17)  %>% gather(key="pheno", value="value", -apple_id) 



#change the phenotype to a factor and reorder the factor
class(abc_pheno_field$pheno)
abc_pheno_field$pheno <- as.factor(abc_pheno_field$pheno)
levels(abc_pheno_field$pheno)

#set the levels for the factors and reorder the levels/rename
abc_pheno_field$pheno <- factor(abc_pheno_field$pheno, levels=c("precocity_16_harv", "flowering_jul_16_harv", "time_ripen_2017", "date_jul_17_harv", "juiciness_16_harv", "weight_avg_17_harv", "firmness_avg_17_harv",  "brix_17_harv", "acidity_17_harv", "tpc", "percent_weight_avg_17", "percent_firmness_avg_17", "percent_brix_17", "percent_acidity_17"))
levels(abc_pheno_field$pheno)


#filter the top ten apples then calculate the max and the median, then summarize to get min and max
top_ten <-  abc_pheno_field %>% filter(apple_id %in% c(48, 205, 160, 167, 1223, 222, 57, 235, 171))
#caluclate min of each trait
min_top_ten <- top_ten %>% group_by(pheno) %>% slice(which.min(value))

#calculate max of each trait
max_top_ten <- top_ten %>% group_by(pheno) %>% slice(which.max(value))

#calculate the median of the top ten
median_top_ten <- top_ten %>% group_by(pheno) %>% summarise(median = median(value, na.rm=T))

#calculate IQR
iqr_top_ten <- top_ten %>% group_by(pheno) %>% summarise(iqr_top_ten= IQR(value, na.rm=T))

#calculate the standard error
error_top_ten <- top_ten %>% group_by(pheno) %>% summarise(error_top_ten= std.error(value, na.rm = T))


#make a table with the max and min values together
min_max_top_ten <- left_join(min_top_ten, max_top_ten, by = "pheno")
min_max_top_ten <- left_join(min_max_top_ten, median_top_ten, by = "pheno")
min_max_top_ten <- left_join(min_max_top_ten, iqr_top_ten, by = "pheno")
min_max_top_ten <- left_join(min_max_top_ten, error_top_ten, by = "pheno")


min_max_top_ten <- min_max_top_ten %>% select(-c(apple_id.x, apple_id.y))

#rename columns
colnames(min_max_top_ten)[colnames(min_max_top_ten)=="value.x"] <- "min_top"
colnames(min_max_top_ten)[colnames(min_max_top_ten)=="value.y"] <- "max_top"
colnames(min_max_top_ten)[colnames(min_max_top_ten)=="median"] <- "median_top"


min_max_top_ten <- min_max_top_ten %>% group_by(pheno)


#########make a table with min and max of the entire collection.

#filter the top ten apples  out then calculate the max and the median, then summarize to get min and max
orchard <- filter(abc_pheno_field, !apple_id %in% c(48, 205, 160, 167, 1223, 222, 57, 235, 171))

#caluclate min of each trait
min_orchard <- orchard %>% group_by(pheno) %>% slice(which.min(value))

#caluclate max of each trait
max_orchard <- orchard %>% group_by(pheno) %>% slice(which.max(value))

#calculate the median of the top ten
median_orchard <- orchard %>% group_by(pheno) %>% summarise(median = median(value, na.rm=T))

#calculate IQR
iqr_orchard <- orchard %>% group_by(pheno) %>% summarise(iqr_orchard= IQR(value, na.rm=T))

#calculate the standard error
error_orchard <- orchard %>% group_by(pheno) %>% summarise(error_top_ten= std.error(value, na.rm = T))

#make a table with the max and min values together
stats_orchard <- left_join(min_orchard, max_orchard, by = "pheno")
stats_orchard <- left_join(stats_orchard, median_orchard, by = "pheno")
stats_orchard <- left_join(stats_orchard, iqr_orchard, by = "pheno")
stats_orchard <- left_join(stats_orchard, error_orchard, by = "pheno")


stats_orchard <- stats_orchard %>% select(-c(apple_id.x, apple_id.y)) %>% rename(min_orchard = value.x) %>% rename(max_orchard = value.y) %>% rename(median_orchard = median)

#combine the top ten and orchard tables

figure_table <- left_join(min_max_top_ten, stats_orchard, by = "pheno")

#########

phenos <- abc_phenos %>% select(apple_id, precocity_16_harv, flowering_jul_16_harv, time_ripen_2017, date_jul_17_harv, juiciness_16_harv, weight_avg_17_harv, firmness_avg_17_harv, brix_17_harv, acidity_17_harv, tpc, percent_weight_avg_17, percent_firmness_avg_17, percent_brix_17, percent_acidity_17)

#make orchard dataframe
all_orchard <- filter(phenos, !apple_id %in% c(48, 205, 160, 167, 1223, 222, 57, 235, 171))
all_orchard <- mutate(all_orchard, apple_type="orchard")
all_orchard <- all_orchard %>% select(-apple_id, precocity_16_harv:apple_type)


orchard_order <- c("apple_type","precocity_16_harv", "flowering_jul_16_harv", "time_ripen_2017", "date_jul_17_harv", "juiciness_16_harv", "weight_avg_17_harv", "firmness_avg_17_harv",  "brix_17_harv", "acidity_17_harv", "tpc", "percent_weight_avg_17", "percent_firmness_avg_17", "percent_brix_17", "percent_acidity_17")
all_orchard <- all_orchard[, orchard_order] 

#make elite dataframe
elite <- filter(phenos, apple_id %in% c(48, 205, 160, 167, 1223, 222, 57, 235, 171))
elite <- mutate(elite, apple_type="elite")
elite <- elite %>% select(-apple_id, precocity_16_harv:apple_type)
elite_order <- c("apple_type","precocity_16_harv", "flowering_jul_16_harv", "time_ripen_2017", "date_jul_17_harv", "juiciness_16_harv", "weight_avg_17_harv", "firmness_avg_17_harv",  "brix_17_harv", "acidity_17_harv", "tpc", "percent_weight_avg_17", "percent_firmness_avg_17", "percent_brix_17", "percent_acidity_17")
elite <- elite[, elite_order] 

comparison_phenos <- full_join(all_orchard, elite)

#calculate wilcoxin p values
precocity_p <- wilcox.test(precocity_16_harv ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
flowering_p <- wilcox.test(flowering_jul_16_harv ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
ripen_p <- wilcox.test(time_ripen_2017 ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
date_p <- wilcox.test(date_jul_17_harv ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
juiciness_p <- wilcox.test(juiciness_16_harv ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
weight_p <- wilcox.test(weight_avg_17_harv ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
firmness_p <- wilcox.test(firmness_avg_17_harv ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
brix_p <- wilcox.test(brix_17_harv ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
acidity_p <- wilcox.test(acidity_17_harv ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
tpc_p <- wilcox.test(tpc ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
change_weight_p <- wilcox.test(percent_weight_avg_17 ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
change_firmness_p <- wilcox.test(percent_firmness_avg_17 ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
change_brix_p <- wilcox.test(percent_brix_17 ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value
change_acidity_p <- wilcox.test(percent_acidity_17 ~ apple_type, data = comparison_phenos,
                   exact = FALSE)$p.value


biglist <- c(precocity_p, flowering_p, ripen_p, date_p, juiciness_p, weight_p, firmness_p,  brix_p, acidity_p, tpc_p, change_weight_p, change_firmness_p, change_brix_p, change_acidity_p)

 
corrected_biglist <- p.adjust(biglist, method = "bonferroni")

#add p-values to the p value column in the figure table.
figure_table$p_values <- biglist
figure_table$p_values_corrected <- corrected_biglist



figure_table <- figure_table %>% group_by(pheno) %>% mutate(prop_dist = (max_top - min_top)/(max_orchard - min_orchard))

figure_table <-figure_table %>% mutate_if(is.numeric, signif, digits = 3)


colnames(figure_table) <- c("Phenotype", "Min 10", "Max 10", "median 10", "IQR 10", "Std. Error 10", "Min", "Max", "Median", "IQR", "Std. Error", "Wilcoxin p-values", "Bonferroni p-values", "Top 10 Proportion of Distribution")


figure_table <- as.data.frame(figure_table)

#decided to take out the IQR columns
figure_table$`IQR 10` <- NULL
figure_table$`IQR` <- NULL


#save table
write.table(figure_table, file = "../figures_tables/supp_table_3.csv", quote = F, sep = ",", row.names = F)



```
#Supp. Table 4: R-square and p-values from 2017 heatmap (reformatted supp_table_4_r2 and supp_table_4_pvals.csv from the Main Data-analyis)

#Supp. Table 5: R-square and p-values from 2016 heatmap (reformatted supp_table_5_r2 and supp_table_5_pvals.csv from the Main Data-analyis)

#Supp. Table 6: Numerical values from the comparison of fruit quality and phenology traits across species, geographic region and use.

#Supp. Table 7: Correlation of year of release with phenotypes.
```{r}

#load data
abc_phenos <- read_csv("../outputs/pheno_meta_data.csv")

abc_phenos <- select(abc_phenos, release_year:tpc, time_ripen_2017)

abc_phenos <- as.matrix(abc_phenos)

r2 = p = c()
for (i in 2:ncol(abc_phenos)) {
  year_correlation <- cor.test(abc_phenos[,"release_year"], abc_phenos[,i], method = "pearson")
  r2[i] = year_correlation$estimate^2
  p[i] = year_correlation$p.val
}
cor_p = p * 38 #38 comparisons to multiple test correct for.
year_vs_phenos = cbind(colnames(abc_phenos), r2, cor_p)
year_vs_phenos = year_vs_phenos[-1,]
year_vs_phenos <- as.data.frame(year_vs_phenos)
year_vs_phenos$cor_p <- as.numeric(as.character(year_vs_phenos$cor_p))
year_vs_phenos <- year_vs_phenos %>% arrange(-desc(cor_p))

#write table to file.
write.table(year_vs_phenos, file = "../figures_tables/supp_tables/supp_table_7.txt", sep = "\t", row.names = F, quote = F)


```



